{\LARGE \textsc{Structural controlability of pollination networks}}

### 2016-05-03 paper/manuscript.Rnw
update methods

\subsection*{Theory of network manageability}

\subsubsection*{Maximum matching}

@@

\subsubsection*{Weighting links}

@@

\subsubsection*{Relative species importance}

\todo[inline]{insert ecological motivation}

@@

We therefore call \textit{driver species} those that are identified as being a superior node in all\todo{check} possible matchings\todo{elaborate concept}, and consequently likely to have a disproportionate effect on the abbundances of other species in the community. Unlike unweighted networks, where a maximum matching is defined by its matching size (the number of matching links), a maximum matching in a weighted network is one whose links have the largest possible total weight among all possible matchings. This means that in weighted networks there is a number of matchings that have maximal size, but are not maximum matchings because their total weight is not maximal.

We therefore identified the driver species in each of the networks by computing all possible matchings in the network that have the same maximal size while allowing for some tolerance regarding the weight. We relax the matching weight requirement because empirical intreaction strengths, from which we calculate dependencies, are stochastic to some extent and depend on the sampling method and intensity. To compute all the maximal matchings, we first generated the network's line graph---an alternative representation in which nodes are the interactions of the original network and links represent common ending points. We then calculated the complement of the line graph---a graph in which nodes are adjacent if and only if they are not adjacent in the line graph. All possible maximum matchings can be found by enumerating all cliques of size $1-n_d$ in this complement graph \autocite{West2001, Csardi2006a}.

\subsection*{Empirical Data}

In the last section we outlined the theoretical and computational underpinings of our approach. Now we illustrate its empirical application using ten pairs of plant-pollinator communities for which we calculated their manageability and the relative importance of species. Each network pair was composed by a community invaded by a plant and a community effectively free of the invasive species (Table S1). The networks were constructed from previously published visitation data collected from pollination communities in Bristol, UK \autocite{Lopezaraiza-Mikel2007} and in Cap de Creus National Park, Spain \autocite{Bartomeus2008}. In the four British uninvaded communities, the non-invaded plots were obtained by experimentally removing all the flowers of the invasive species \textit{Impatients grandulifera}. In contrast, the Spanish uninvaded communities were obtained from plots that had not yet been colonised by either of the invasive species \textit{Carpobrotus affine acinaciformis} or \textit{Opuntia stricta}.

The structure of these networks was determined using visitation frequency which has been shown to be an appropriate surrogate for inter-specific effects in pollination networks \autocite{Vazquez2005, Bascompte2006}. To examine the importance of this decision, we also evaluated the effect of using pollinator efficiency or importance as alternative measures of species interactions \autocite{NeEman2010, Ballantyne2015}, and we found consistent results for all of these options(see Supplementary Information).

We started by quantifying the manageability of the analysed pollination communities. To do that, we calculated the maximum matching of the corresponding pollination network, and estimated the minimum number of species that need directed intervecntions to fully control the species abbundances in the community. If a network had more than one component (two species are in different components if there exist no path between them) we analysed the largest one. In order to check that

We then performed a paired Spearman correlation test to compare the manageability obtained when including bidirectional effects between interacting species, and when simplyfing the network to include only the dominant direction of the asymmetry.

Naively one might expect to see a negative relationship between diversity and manageability \autocite{Menge1995}. We therefore used a set of generalised linear models to investigate the relatioship between community manageability and (i) the total number of species, (ii) plant richness, (iii) polinators richness, and (iv) the ratio of plants to pollinators richness. Because both invaded and British communities tended to be more diverse, we also assesed whether observed differences in manageability were due to these two factors. Competing models were assessed by comparing the Akaike Information Criterion corrected for small sample sizes (AICc).

To quantify the extent to which manageability is characteristic of the network structure\todo{failed to come up with a better wording}, we compared the manageability of empirical networks to networks generated by two null models of visitation. One set of random visitation networks was generated by maintaining the species degree (the number of interactions per species) while varying the number of visits per species. A second set of random network mantained the species strength (total sum of visits) while varying the degree of species \autocite{Oksansen2016}. We then calculated the mutual dependences and interaction asymmetries of the random networks and calculated the p-value of the manageability of each empirical network and the corresponding manageability of 999 randomisations.

We also quantified the extent to which the relative manageability of a pollination community is given by the dominant direction of control, we mantained the structure of network but randomised the direction of the interaction asymmetries. This is, we first calculated the assymetries for each community and then shuffled the direction of the link between species. Similar as in previouls null models, we calculated the p-value of the empirical manageability and that of the randomisations.

 However the weight of species dependencies is unexact, a reflection of the fact that interaction strengths and sampling methods themselves are stochastic to some extent.

We therefore identified the driver species in each of the networks by computing all possible matchings in the network that have the same size and have a weight above a cut-off of 0.5 times the weight of the maximum matching\todo{justify}.

Because of the number of matchings depends on the cut-off and the weighting of the links, we evaluated the effects of i) using a different cut-offs value between 0 and 1 and ii) weighting the links by the number of visits.

Similar as when we compared the manageability obtained when assuming bidirectional links or single links determined by the asymmetry (\autoref{fig:explain_cd}), we compared the realtive importance of species under these two assumptions by calculating a paired Spearman correlation test between species' relative importance. Nevertheless because we are more interested on the agreement between highly ranked species we also calculated a weighted correlation coefficient $r_{w2}$ which gives more weight to higer ranks (\autoref{PintoDaCosta2015}), and the Jaccard similarity of the driver species identified by both methods.

We then tackled the question whether some species level structural properties can predict the relative importance of driver species. Here, we used a Generalised Linear Mixed-Effects model (\autoref{Bates2015}) to evaluate the effect on the species' relative importance of measures of centrality (degree, betweenness, and overlap), measures related to network robustness (contribution to nestedness), and measures of strength of association and depedency (visitation and dependency strength) \autocite{Dormann2008}. The community was included as a random effect. Candidate models were compared using AIC and the relative importance of the explanatory variables was evaluated using the sum of Akaike weights over candidate models that accounted for 95\% of the evidence \autocite{Burnham2003, Bates2014, Barton2016}.

>>>>

\section*{Theory of community manageability}

\subsection*{Maximum matching}

@@

\subsection*{Weighting links}

@@

\subsection*{Relative species importance}

\todo[inline]{insert ecological motivation?}

@@

We therefore call \textit{driver species} those that are identified as being a superior node in all\todo{check}  possible matchings, and consequently likely to have a disproportionate effect on the abbundances of other species in the community. Unlike unweighted networks, where a maximum matching is defined by its matching size (the number of matching links), a maximum matching in a weighted network is one whose links have the largest possible total weight among all possible matchings. This means that in weighted networks there is a number of matchings that have maximal size, but are not maximum matchings because their total weight is not maximal.

Consequently, we identified the driver species in each of the networks by computing all possible matchings in the network that have the same maximal size while allowing for some tolerance regarding the weight. We relax the matching weight requirement because empirical intreaction strengths, from which we calculate dependencies, are stochastic to some extent and depend on the sampling method and intensity. To compute all the matchings with maximal size, we first generated the network's line graph---an alternative representation in which nodes are the interactions of the original network and links represent common ending points. We then calculated the complement of the line graph---a graph in which nodes are adjacent if and only if they are not adjacent in the line graph. All possible maximum matchings can be found by enumerating all cliques of size $1-n_d$ in this complement graph \autocite{West2001, Csardi2006a}.

In the last section we outlined the theoretical and computational underpinings of our approach. Now we illustrate its empirical application using ten pairs of plant-pollinator communities for which we calculated their manageability and the relative importance of species. Each network pair was composed by a community invaded by a plant and a community effectively free of the invasive species (Table S1). The networks were constructed from previously published visitation data collected from pollination communities in Bristol, UK \autocite{Lopezaraiza-Mikel2007} and in Cap de Creus National Park, Spain \autocite{Bartomeus2008}. In the four British uninvaded communities, the non-invaded plots were obtained by experimentally removing all the flowers of the invasive species \textit{Impatients grandulifera}. In contrast, the Spanish uninvaded communities were obtained from plots that had not yet been colonised by either of the invasive species \textit{Carpobrotus affine acinaciformis} or \textit{Opuntia stricta}. The structure of these networks was determined using visitation frequency which has been shown to be an appropriate surrogate for inter-specific effects in pollination networks \autocite{Vazquez2005, Bascompte2006}. To examine the importance of this decision, we also evaluated the effect of using pollinator efficiency or importance as alternative measures of species interactions \autocite{NeEman2010, Ballantyne2015}, and we found consistent results for all of these options(see Supplementary Information).

\subsection*{Community manageability}

We started by quantifying the manageability of the analysed pollination communities. To do that, we calculated the maximum matching of the corresponding pollination network, and estimated the minimum number of species that need directed intervecntions to fully control the species abbundances in the community. If a network had more than one component (two species are in different components if there exist no path between them) we analysed the largest one. Although understanding the distribution of manageability across ecological networks is an useful result in itself, we wanted to test wheter perturbations, like invasion status, have an impact on manageability values. As, naively one might expect to see a negative relationship between diversity and manageability \autocite{Menge1995}, we used a set of generalised linear models to investigate the effect of invasion status while controling for \textit{(i)} the total number of species, \textit{(ii)} plant richness, \textit{(iii)} polinators richness, \textit{(iv)} the ratio of plants to pollinators richness, and \textit{(v)} the study site. Competing models were assessed by comparing the Akaike Information Criterion corrected for small sample sizes (AICc).

Subsequently we explored the extent to which structural factors in the empirical networks influence their manageability. Previous studies found a direct link between a network's degree distribution and the number of nodes necessary to fully control it \autocite{Liu2011}, but this relationship has not been explored in weighted ecological networks. We therefore compared the manageability of the empirical networks to networks generated by three null models. One set of random visitation networks was generated by maintaining each species' degree (its number of interactions) while varying the number of visits per species. A second set of random networks mantained the species' strength (its total sum of visits) while varying their degrees. Randomisations were generated using vegan 2.3-3 \autocite{Oksansen2016}. Finally a three set of fully random networks maintained only the connectivity of the empirical network, but both degree and weight were allowed to vary. We then calculated the mutual dependences and interaction asymmetries of the random networks and calculated their manageability using the maximum matching algorithm. Finally we calculated the p-value of the manageability of each empirical network and the corresponding manageability of 999 randomisations.

The dependency asymetry plays a fundamental role in determining the direction of control in a two species interaction and therefore has the potential to structure the network matching. We therefore performed an additional randomisation in which we mantained the structure of network but randomised the direction of the interaction asymmetries. This is, we first calculated the assymetries for each community and then shuffled the direction of the link between species. Similar as in previouls null models, we calculated the p-value of the empirical manageability compared to that of the randomisations.

\subsection*{Relative importance of species}

Our third question is related to how different species differ on their importance for the population dynamics of the community. To quantify this importance we computed all possible matchings that have the same size as the maximum matchings and in each network. We then calculated the frequency to which each species is deemed to be a superior node in matchings with a weight equal or larger to 0.5 times the weight of the maximum matchings. We choose this threshold as it provided a high agreement between the relative importance of species calculated by weighting interactions with the mutial dependencies or the depndence asymmetry as well as between the network is quantified by visitation or pollination efficiency. Nevertheless, the choice of the threshold only had a negligible impact on the results (Supplementary information).

The number of maximal matchings increasses exponentially with the number of links and nodes in a network. When we use mutual dependecies, the large number of links in the most diverse communities renders the computation unfeasible. Taking this limitation into account we compared the relative network manageability obtained when using assimetries or when using mutual dependencies for eleven networks---for which the number of matchings was smaller than $4\times 10^{10}$. Results from a paired Spearman correlation test, the weighted correlation coefficient $r_{w2}$---which gives more weight to higer ranks \autocite{PintoDaCosta2015}, and the Jaccard similarity of the driver species in both approaches, suggested a high similarity in relative ranks. Therefore, we used the results from the assymetries, rather than mutiual dependences, to further investigate the factors that determine species' relative importance.

We then asked the question of whether some species-level structural properties can predict the relative importance of species. Here, we used a Generalised Linear Mixed-Effects model to evaluate the effect on the species' relative importance of a suite of structural measures. We included measures of centrality (degree and eigen-centrality) which have been found to be strong predictors of importance in a coextintion context\todo{add ref}; measures related to network robustness (contribution to nestedness) as nestedness has been proposed as one of the key properties that promote stability in mutualistic networks \autocite{Saavedra2011}; and measures of strength of association and depedency (visitation and dependency strength) as their skewed distribution is characteristic of plant-pollinator networks and determine the extent of interspecific effects. All metrics were calculated using bipartite 2.06 \autocite{Dormann2008}. To allow for variation among different ecological communities, this factor was included as a random effect \autocite{Bates2015a}. Candidate models were compared using AIC and the relative importance of the explanatory variables was evaluated using the sum of Akaike weights over candidate models that accounted for 95\% of the evidence \autocite{Burnham2003, Bates2014, Barton2016}.

Invasive species have been shown to have disproportionate effects on the communities they invade. So we examined the extent to which invasive species tend to be driver species---species whith the maximum importance. As our invasive species were plants, we compared the relative importance of invasive species to those of other plants in the community and to which degree they differ when accounting for the structural factors that determine their importance\todo{needs to be developed once the results section is rewritten}.
